---
title: "Final report"
author: "Pengfei Lou"
date: "`r Sys.Date()`"
output: html_document
---

```{r,warning=FALSE}
pacman::p_load(
  ggplot2,
  tibble,
  dplyr,
  lmtest,
  knitr,
  gtsummary,
  caret,
  pROC,
  MASS,
  mice,
  car,
  stats,
  sjPlot,
  here
)
```

```{r, echo = FALSE}

opts_chunk$set(
  echo=TRUE,
  warning=FALSE,
  message=FALSE
)

```

# Aim 1

## Deal with missing value

```{r}
here::i_am(
  "507.Rmd"
)

file_path<-here::here("data/pregnancy.csv")
data<-read.csv(file_path,header = TRUE)
na_counts<-colSums(is.na(data))
print(na_counts)
```

```{r}
data$mstatus[is.na(data$mstatus)]<-6
data$mstatus<-factor(data$mstatus,levels = c("0","1","2","3","4","5","6"))

data<-mice(data,m=5,method = "pmm",seed = 123)

#plot(data)

data <- complete(data, 1)
na_counts<-colSums(is.na(data))
print(na_counts)
```

The missing value mainly exist in the mstatus variable representing the meaning of 'don't know'. Thus, NA in mstatus shouldn't be deleted. Instead, "6" is assigned to replace the 'don't know'. The missing value in the wt is inserted by mean value.


## EDA Analysis

```{r}
data <- data %>%
  mutate(
    hb_dl = hb / 10,
    anemia = ifelse(hb_dl < 12, 1, 0),
    BMI = wt / (ht / 100)^2
  ) %>%
  dplyr::select(-ht, -wt, -hb, -hb_dl) 

table<-data%>%
  tbl_summary(by = anemia)
table
```

The summary table is divided by No anemia(0) and anemia(1). There is imbalance existing in outcome(Aim 3). Thus, upsampling is used.

## Data scaling

```{r}
boxplot(data, main = "Boxplot for Identifying Outliers")
```


```{r}
data$crp <- log(data$crp+1)
data$sf <- log(data$sf+1)
data$sfo <- log(data$sfo+1)
data$si<-log(data$si+1)
data$vitb12<-log(data$vitb12+1)
data$age_yr<-log(data$age_yr+1)
data$gagewks<-log(data$gagewks+1)
data$BMI<-log(data$BMI+1)
```


## Upsampling

```{r}
data_anemia<-data%>%
  filter(anemia==1)
data_no_anemia<-data%>%
  filter(anemia==0)

times<-nrow(data_no_anemia)/nrow(data_anemia)

data_anemia_upsampled<-data_anemia[rep(1:nrow(data_anemia),each=ceiling(times)),]

data_balanced<-rbind(data_anemia_upsampled,data_no_anemia)

table(data_balanced$anemia)

summary(data)
summary(data_balanced)
```



```{r}

micronutrients <- c("sf", "sfo", "si", "vitb12")

for(micronutrient in micronutrients) {
  model <- lm(as.formula(paste(micronutrient, "~ crp")), data = data)
  cat("\n","Analyzing", micronutrient, "\n")
  term <- c("crp")
  for(term in term) {
    if (summary(model)$coefficients[term, "Pr(>|t|)"] < 0.05){
      print(summary(model))
      cat(term, "is significant for", micronutrient, "\n")
    }
    else{
      cat(term,"is not significant for",micronutrient, "\n")
      }
  }
}

```

CRP is significant for si and vitb12.


```{r}
data$trimester <- as.factor(ifelse(data$trimester == "4", "3", data$trimester))
data$trimester <- factor(data$trimester)  
```

The  trimester variable in level 4 only has one observation which is not statistically enough to make estimate. Thus, this level is combined with the level 3.

```{r}

# Define predictor variables and covariates
micronutrients <- c("sf", "sfo", "si", "vitb12")  # Dependent variables
covariates <- c("age_yr", "gagewks", "trimester")  # Independent variables
predictor <- "crp"  # Main independent variable

# Function to construct model formulae
build_formula <- function(response, predictor, covars) {
  paste(response, "~", predictor, "+", paste(covars, collapse = " + "))
}

# Initialize a list to store models
model_results <- list()

# Fit models for each micronutrient with varying covariate combinations
for (micro in micronutrients) {
  covariate_combinations <- unlist(lapply(1:length(covariates), function(n) combn(covariates, n, simplify = FALSE)), recursive = FALSE)
  for (cov_combination in covariate_combinations) {
    formula <- build_formula(micro, predictor, cov_combination)
    model <- try(lm(formula, data = data), silent = TRUE)
    if (!inherits(model, "try-error") && "lm" %in% class(model)) {
      model_results[[formula]] <- model  # Store the model itself
    } else {
      warning(paste("Model fitting failed or model not valid for formula:", formula))
      model_results[[formula]] <- NULL
    }
  }
}

# Print results and conclusions for each model
for (formula in names(model_results)) {
  cat("\nModel:", formula, "\n")
  model <- model_results[[formula]]
  if (!is.null(model) && "lm" %in% class(model)) {
    result <- summary(model)
    print(result$coefficients)
    cat("Residual Standard Error:", result$sigma, "\n")
    cat("Multiple R-squared:", result$r.squared, "\n")
    cat("Adjusted R-squared:", result$adj.r.squared, "\n")
    cat("AIC:", AIC(model), "\n")  # Calculate AIC using the model object

    # Conclusions based on p-values and R-squared
    significant_predictors <- summary(model)$coefficients[,4] < 0.05
    if (any(significant_predictors)) {
      cat("Conclusion: The model shows significant predictors.\n")
    } else {
      cat("Conclusion: No significant predictors found.\n")
    }
  } else {
    cat("Model fitting failed or returned NULL.\n")
  }
}

```

sf~crp: The existence of maternal age and/or gestational age (or pregnancy trimester) has a positive or negative contribution to the association.

sfo~crp: The existence of maternal age and/or gestational age (or pregnancy trimester) has a  positive or negative contribution to the association.

si~crp: The existence of maternal age and/or gestational age (or pregnancy trimester) has a  positive or negative contribution to the association.

vitb12~crp: The existence of maternal age and/or gestational age (or pregnancy trimester)  has a  positive or negative contribution to the association.

It seems like the trimester variable in level 2 and 3 is statistical significantly.

## Two way interaction

### One interaction

```{r}
data$trimester <- factor(data$trimester)
levels(data$trimester)  # Check the levels

```


```{r}
micronutrients <- c("sf", "sfo", "si", "vitb12")

# Function to fit model and check significance
fit_and_evaluate_model <- function(formula, data, interaction_term = NULL) {
  model <- tryCatch({
    lm(as.formula(formula), data = data)
  }, error = function(e) {
    cat("Error fitting model:", e$message, "\n")
    return(NULL)
  })

  if (!is.null(model)) {
    summary_model <- summary(model)
    # Check if the interaction term exists and is estimable
    if (!is.null(interaction_term) && interaction_term %in% rownames(summary_model$coefficients)) {
      p_value <- summary_model$coefficients[interaction_term, "Pr(>|t|)"]
      significant <- p_value < 0.05
      cat(sprintf("%s interaction with p-value = %f is %s\n", interaction_term, p_value, ifelse(significant, "significant", "not significant")))
    } else if (!is.null(interaction_term)) {
      cat(sprintf("Interaction term '%s' not found or not estimable in the model.\n", interaction_term))
    }
  }
}

# Main analysis loop
for (micronutrient in micronutrients) {
  cat("\nAnalyzing", micronutrient, "\n")
  base_formula <- paste(micronutrient, "~ crp + age_yr + gagewks +trimester")
  interaction_terms <- c("crp:age_yr", "crp:gagewks")

  # Fit and evaluate the base model
  fit_and_evaluate_model(base_formula, data)

  # Fit and evaluate models with interactions
  for (interaction_term in interaction_terms) {
    extended_formula <- paste(base_formula, "+", interaction_term)
    fit_and_evaluate_model(extended_formula, data, interaction_term)
  }
}


```
Since the trimester is factor variable, I have to divide this part into two for result clarification.


```{r}

micronutrients <- c("sf", "sfo", "si", "vitb12")

# Function to fit models and evaluate interactions dynamically
fit_and_evaluate_model <- function(formula, data) {
    model <- tryCatch({
        lm(as.formula(formula), data = data)
    }, error = function(e) {
        cat("Error fitting model:", e$message, "\n")
        return(NULL)
    })

    if (!is.null(model)) {
        summary_model <- summary(model)
        coefficients_summary <- summary_model$coefficients
        # Dynamically find and test interaction terms involving 'crp' and 'trimester'
        interaction_terms <- grep("crp:trimester", rownames(coefficients_summary), value = TRUE)
        
        if (length(interaction_terms) > 0) {
            for (term in interaction_terms) {
                p_value <- coefficients_summary[term, "Pr(>|t|)"]
                significant <- p_value < 0.05
                cat(sprintf("%s interaction with p-value = %f is %s\n", term, p_value, ifelse(significant, "significant", "not significant")))
            }
        } else {
            cat("No interaction terms involving 'crp' and 'trimester' found or estimable in the model.\n")
        }
    }
}

# Main analysis loop
for (micronutrient in micronutrients) {
    cat("\nAnalyzing", micronutrient, "\n")
    # The base formula includes interactions with all levels of 'trimester' that R automatically creates
    base_formula <- paste(micronutrient, "~ crp + age_yr + gagewks + trimester + crp:trimester")
    
    # Evaluate the model with the base formula and interactions
    fit_and_evaluate_model(base_formula, data)
}

```


### Two interactions

```{r,warning=FALSE}
# Define the base predictor and covariates
predictor <- "crp"
covariates <- c("age_yr", "gagewks", "trimester")
micronutrients <- c("sf", "sfo", "si", "vitb12")

# Create all possible pairs of interaction terms
interaction_terms <- c(paste(predictor, covariates[1], sep=":"), paste(predictor, covariates[2], sep=":"), paste(predictor, covariates[3], sep=":"))
interaction_pairs <- combn(interaction_terms, 2, simplify = FALSE)

# Build and evaluate models for each micronutrient and each pair of interaction terms
results <- list()
for (micronutrient in micronutrients) {
    results[[micronutrient]] <- list()
    cat("\nAnalyzing", micronutrient, "\n")
    
    for (pair in interaction_pairs) {
        interaction_formula <- paste(micronutrient, "~", predictor, "+", paste(covariates, collapse = "+"), "+", paste(pair, collapse = "+"))
        model <- lm(as.formula(interaction_formula), data = data)
        summary_model <- summary(model)
        
        # Store model summary
        results[[micronutrient]][paste(pair, collapse=" & ")] <- summary_model
        
        # Output the model's summary
        cat("\nModel with interactions", paste(pair, collapse=" and "), ":\n")
        print(summary_model)
        
        # Check if interactions are significant
        p_values <- sapply(pair, function(x) summary_model$coefficients[x, "Pr(>|t|)"])
        significant <- all(p_values < 0.05)
        if (significant) {
            cat("Both interactions", paste(pair, collapse=" and "), "are significant for", micronutrient, "\n")
        } else {
            cat("No interactions are significant for", micronutrient, "\n")
        }
    }
}

```

Both interactions crp:gagewks and crp:trimester are significant for vitb12.

### Three interactions

```{r}

micronutrients <- c("sf", "sfo", "si", "vitb12")

for(micronutrient in micronutrients) {
  cat("Analyzing", micronutrient, "\n")
  model_no_interaction <- lm(as.formula(paste(micronutrient, "~ crp+age_yr+trimester+gagewks")), data = data)
  model_interaction <- lm(as.formula(paste(micronutrient, "~ crp + age_yr + gagewks + trimester + crp:age_yr + crp:gagewks + crp:trimester")), data = data)
  interaction_terms <- c("crp:age_yr", "crp:gagewks","crp:trimester")
  for(term in interaction_terms) {
    if (summary(model_interaction)$coefficients[term, "Pr(>|t|)"] < 0.05){
      print(summary(model_interaction))
      cat(term, "interaction is significant for", micronutrient, "\n")
      
    }
    else{
      cat(term,"interaction is not significant for",micronutrient, "\n")
      }
      
    }
  }
#interaction_plot <- ggplot(data_scaled, aes(x=crp, y=sfo, color=factor(gagewks),group=factor(gagewks))) +
                #geom_point() +
                #geom_smooth(method="lm", se=TRUE) +
                #labs(title=paste("Interaction of CRP and gagewks on sfo"))
            #print(interaction_plot)
            
#interaction_plot <- ggplot(data_scaled, aes(x=crp, y=sfo, color=factor(trimester),group=factor(trimester))) +
                #geom_point() +
                #geom_smooth(method="lm", se=TRUE) +
                #labs(title=paste("Interaction of CRP and trimester on sfo"))
            #print(interaction_plot)


```

After considering the interaction among maternal age, gestational age, pregnancy trimester and CRP, only the interaction among CRP,pregnancy trimester and gestational age is significant for vitb12 when three interactions all exist.

### Explore more interations among age_yr, gagewks and trimester

```{r}
micronutrients <- c("sf", "sfo", "si", "vitb12")

for(micronutrient in micronutrients) {
    cat("Analyzing", micronutrient, "\n")
    model_interaction <- lm(as.formula(paste(micronutrient, "~ crp + age_yr + gagewks + trimester + age_yr:crp + age_yr:gagewks + age_yr:trimester")),data=data)
    summary_interaction<-summary(model_interaction)
    interaction_terms <- c("crp:age_yr", "age_yr:gagewks", "age_yr:trimester")
    significant_found <- FALSE

    for(term in interaction_terms) {
        p_value <- summary(model_interaction)$coefficients[term, "Pr(>|t|)"]
        if (!is.na(p_value) && p_value < 0.05) {
            if (!significant_found) {  # Print the model summary only once if there's at least one significant interaction
                print(summary_interaction)
                significant_found <- TRUE
            }
            cat(term, "interaction is significant for", micronutrient, "\n")
        } else {
            cat(term, "interaction is not significant for", micronutrient, "\n")
        }
    }
  
    cat("\n") 
}


```


```{r}

data$age_group <- cut(data$age_yr,
                      breaks = c(15, 20, 30, 41), 
                      labels = c("16-20", "21-30", "31-41"),
                      include.lowest = TRUE)
# List of micronutrient variables to iterate over
micronutrients <- c("sf", "sfo", "si", "vitb12")

# Loop through each micronutrient
for(micronutrient in micronutrients) {
    cat("Analyzing", micronutrient, "\n")
  
    # Model with interaction
    model_interaction <- lm(as.formula(paste(micronutrient, "~ crp + age_yr + gagewks + trimester + gagewks:crp + gagewks:age_yr + gagewks:trimester")),data=data)
    summary_interaction<-summary(model_interaction)
    # Check for significant interactions and visualize if necessary
    interaction_terms <- c("crp:gagewks", "age_yr:gagewks", "gagewks:trimester")
    significant_found <- FALSE

    for(term in interaction_terms) {
        p_value <- summary(model_interaction)$coefficients[term, "Pr(>|t|)"]
        if (!is.na(p_value) && p_value < 0.05) {
            if (!significant_found) {  # Print the model summary only once if there's at least one significant interaction
                print(summary_interaction)
                significant_found <- TRUE
            }
            cat(term, "interaction is significant for", micronutrient, "\n")
            
        } else {
            cat(term, "interaction is not significant for", micronutrient, "\n")
        }
    }
  
    cat("\n") 
}

#interaction_plot <- ggplot(data_scaled, aes(x=gagewks, y=sf, color=factor(age_group),group=factor(age_group))) +
                #geom_point() +
                #geom_smooth(method="lm", se=TRUE) +
                #labs(title=paste("Interaction of age and gagewks on sf"))
            #print(interaction_plot)
```
gagewks:trimester interaction is significant for sf and vitb12.

```{r}
# List of micronutrient variables to iterate over
micronutrients <- c("sf", "sfo", "si", "vitb12")

# Loop through each micronutrient
for(micronutrient in micronutrients) {
    cat("Analyzing", micronutrient, "\n")
  
    # Model with interaction
    model_interaction <- lm(as.formula(paste(micronutrient, "~ crp + age_yr + gagewks + trimester + trimester:crp + trimester:age_yr + trimester:gagewks")),data=data)
    
    # Check for significant interactions and visualize if necessary
    interaction_terms <- c("crp:trimester", "age_yr:trimester", "gagewks:trimester")
    significant_found <- FALSE

    for(term in interaction_terms) {
        p_value <- summary(model_interaction)$coefficients[term, "Pr(>|t|)"]
        if (!is.na(p_value) && p_value < 0.05) {
            if (!significant_found) {  # Print the model summary only once if there's at least one significant interaction
                print(summary_interaction)
                significant_found <- TRUE
            }
            cat(term, "interaction is significant for", micronutrient, "\n")
            
           
        } else {
            cat(term, "interaction is not significant for", micronutrient, "\n")
        }
    }
  
    cat("\n") 
}

```
gagewks:trimester interaction is significant for sf and vitb12

```{r}
micronutrients <- c("sf", "sfo", "si", "vitb12")
for(micronutrient in micronutrients) {
  cat("Analyzing", micronutrient, "\n")
  
  formula_interaction <- as.formula(paste(micronutrient, "~ crp * age_yr * gagewks * trimester"))
  model_interaction <- lm(formula_interaction, data = data)

  model_summary <- summary(model_interaction)

  interaction_terms <- names(coef(model_summary))[grepl(":", names(coef(model_summary)))]
  
  significant_interactions <- c()
  
  for(term in interaction_terms) {
    p_value <- coef(model_summary)[term, "Pr(>|t|)"]
    if (!is.na(p_value) && p_value < 0.05) {
      significant_interactions <- c(significant_interactions, term)
      cat(term, "interaction is significant for", micronutrient, "\n")
    }
  }
  
  if (length(significant_interactions) > 0) {
    print(model_summary)
    for(term in significant_interactions) {

      vars <- unlist(strsplit(term, ":"))
    }
  } else {
    cat("No significant interactions found for", micronutrient, "\n")
  }
  
  cat("\n") 
}

```
If conducting high level interaction like three way, four way interaction will lead to no significant interactions. Maybe machine fail to detect the difference due to the complicated conditions.



## Model diagnostics

```{r}
model<-lm(si~crp + age_yr + gagewks + trimester  + crp:gagewks + crp:trimester+crp:age_yr,data=data)
summary(model)
res_fit_data<-tibble(
  residuals=model$residuals,
  fitted=model$fitted.values
)

res_fit_plot<- res_fit_data %>%
  ggplot(aes(x=fitted,y=residuals))+
  geom_point()+
  geom_hline(yintercept = 0,color="red")
res_fit_plot
dwtest(model)

qqnorm(residuals(model))
qqline(residuals(model))

shapiro.test(residuals(model))

```

```{r}
diffits_model <- dffits(model)
cd_model <- cooks.distance(model)
dfbetas_model <- dfbetas(model)

combined_measures <- cbind(diffits_model, cd_model, dfbetas_model)

coefficient_names <- names(coef(model))
dfbetas_names <- paste0("dfbetas_", coefficient_names)

colnames(combined_measures) <- c("DFFITS", "Cooks Dist", dfbetas_names)

#combined_measures
```

```{r}
cd_model <- cooks.distance(model)

# Plot Cook's distances
plot(cd_model, type="h", main="Cook's distance", ylab="Cook's distance", xlab="Index")
abline(h = 4/(nrow(data)-length(model$coefficients)), col="red")  # Rule-of-thumb threshold line


```


```{r}
vif_values<-vif(model)
print(vif_values)

```


**Linearity**: There is no systematic pattern in the Residual vs Fitted plot. No nonlinearity will be found.

**Homoscedasticity**: There is no systematic pattern in the Residual vs Fitted plot. No homoscedasticity will be found.

**Independence of errors**: The Residual vs Fitted plot and Durbin-Watson (DW) test show that the assumption of independence of errors is met.

The Durbin-Watson (DW) test gives a DW statistic of  1.4214 with a p-value of 4.2e-14. 

The DW statistic ranges from 0 to 4, where:
  - A value around 2 suggests no autocorrelation.
  - A value significantly less than 2 suggests positive autocorrelation.
  - A value significantly greater than 2 suggests negative autocorrelation.

The DW statistic  is less than 2, which suggests the presence of positive autocorrelation.

A low p-value indicates that there is  enough evidence to reject the null hypothesis of no autocorrelation. 

In summary, based on the Durbin-Watson test result,  there is  significant autocorrelation in the residuals of the model. This suggests that the assumption of independence of errors is not met.

**Normality of residuals**: The QQplot shows most of the point fall on a 45 degree line except for both ends. Thus,Shapiro-Wilk normality test is used to see further.

The value of W = 0.90177 indicates the test statistic from the Shapiro-Wilk test. The value which is more closer to 1 indicates  how close the residuals are to the normal distribution. A value of 1 would indicate a perfect normal distribution.
  
The p-value < 2.2e-16 indicates the null hypothesis (that the data is normally distributed) is not true.

Therefore, the Q-Q plot alongside the Shapiro-Wilk test result show the residuals is not normally distributed.


**Multicollinearity**: VIF function shows there is  multicollinearity.

**DFFITS,Cook's Distance,DFBETAS**: There are no potentially problematic observations to determine if they are exerting on overall strong influence on the fitted regression model.

I conduct a overall t test showing the pvalue < 0.05 which means the model is statistically significant.


# Aim 2

```{r,echo=FALSE,include=FALSE}
here::i_am(
  "507.Rmd"
)

file_path<-here::here("data/pregnancy.csv")
data<-read.csv(file_path,header = TRUE)

data$mstatus[is.na(data$mstatus)]<-6
data$mstatus<-factor(data$mstatus,levels = c("0","1","2","3","4","5","6"))

data<-mice(data,m=5,method = "pmm",seed = 123)

data <- complete(data, 1)
data_balanced$mstatus<-factor(data_balanced$mstatus)
data_balanced$tpreg<-factor(data_balanced$tpreg)

data <- data %>%
  mutate(
    hb_dl = hb / 10,
    anemia = ifelse(hb_dl < 12, 1, 0),
    BMI = wt / (ht / 100)^2
  ) %>%
  dplyr::select(-ht, -wt, -hb, -hb_dl) 

data$crp <- log(data$crp+1)
data$sf <- log(data$sf+1)
data$sfo <- log(data$sfo+1)
data$si<-log(data$si+1)
data$vitb12<-log(data$vitb12+1)
data$age_yr<-log(data$age_yr+1)
data$gagewks<-log(data$gagewks+1)
data$BMI<-log(data$BMI+1)

```

## CRP and  micronutrients 
```{r}

# Ensure the outcome variable 'anemia' is a factor with appropriate levels if it's binary
data$anemia <- factor(data$anemia, levels = c(0, 1))

# Define the array of micronutrients that you are interested in analyzing
micronutrients <- c("sf", "sfo", "si", "vitb12")

# Function to build a logistic regression model for a given micronutrient
build_model <- function(micronutrient, data_balanced) {
    formula <- as.formula(paste("anemia ~ crp+", micronutrient))
    model <- glm(formula, family = binomial(link = "logit"), data = data_balanced)
    return(model)
}

# Initialize a list to store the model summaries
results <- list()
model_summaries <- list()

# Loop through each micronutrient to build models and collect summaries
for (micronutrient in micronutrients) {
    cat("\nAnalyzing", micronutrient, "combined with CRP:\n")
    model <- build_model(micronutrient, data_balanced)
    summary_model <- summary(model)
    results[[micronutrient]] <- model
    model_summaries[[micronutrient]] <- summary_model
    
    # Print the summary to the console
    print(summary_model)
    
    # Optionally, show only the coefficients part
    cat("\nCoefficients for", micronutrient, ":\n")
    print(coef(summary_model))
}

# Comparing models based on AIC values
aic_values <- sapply(results, AIC)
cat("\nAIC values for models:\n")
print(aic_values)


for (micronutrient in micronutrients) {
    coef_df <- as.data.frame(coef(summary(results[[micronutrient]])))
    plot <- ggplot(coef_df, aes(x = reorder(rownames(coef_df), Estimate), y = Estimate, fill = Estimate < 0)) +
            geom_bar(stat = "identity") +
            coord_flip() +
            labs(title = paste("Effect sizes for", micronutrient), x = "Terms", y = "Coefficient Estimate") +
            theme_minimal()
    print(plot) 
}

```

## Interaction
```{r}
data$anemia <- factor(data$anemia, levels = c(0, 1))
# Define the array of micronutrients and other covariates
micronutrients <- c("sf", "sfo", "si", "vitb12")
covariates <- c("age_yr", "BMI", "trimester", "gagewks")  # Clinical and demographic covariates

# Function to generate model formulas including interactions
generate_formula <- function(micronutrient) {
    main_effects <- paste(c("crp", micronutrient, covariates), collapse = " + ")
    interactions <- paste("crp", c(micronutrient, covariates), sep=":", collapse = " + ")
    full_formula <- paste("anemia ~", main_effects, "+", interactions)
    return(as.formula(full_formula))
}

# Loop through each micronutrient, build models, and store results
results <- list()

for (micronutrient in micronutrients) {
    cat("\nAnalyzing", micronutrient, "\n")
    
    # Generate the formula for the current micronutrient
    formula <- generate_formula(micronutrient)
    
    # Build the logistic regression model using binomial family
    model <- glm(formula, family = binomial(link = "logit"), data = data)
    
    # Store the model summary
    results[[micronutrient]] <- summary(model)
    
    # Print the summary to the console
    print(results[[micronutrient]])
}

# Evaluate and display significant interactions
for (micronutrient in micronutrients) {
    cat("\nResults for", micronutrient, ":\n")
    coef_summary <- coef(results[[micronutrient]])
    significant_interactions <- coef_summary[grepl("crp:", rownames(coef_summary)), "Pr(>|z|)"] < 0.05
    
    if (any(significant_interactions)) {
        cat("Significant interactions found for", micronutrient, ":\n")
        print(coef_summary[significant_interactions, , drop = FALSE])
    } else {
        cat("No significant interactions found for", micronutrient, "\n")
    }
}

# Visualize significant interactions if any
for (micronutrient in micronutrients) {
    if (any(results[[micronutrient]]$coefficients[grepl("crp:", names(results[[micronutrient]]$coefficients)), "Pr(>|z|)"] < 0.05)) {
        ggplot(data, aes_string(x = "crp", y = micronutrient, color = "factor(anemia)")) +
        geom_point(alpha = 0.5) +
        geom_smooth(method = "glm", method.args = list(family = "binomial"), formula = results[[micronutrient]]$call$formula, se = FALSE) +
        labs(title = paste("Effect of CRP and", micronutrient, "on Anemia"))
    }
}


```

# Aim 3


```{r,echo=FALSE,include=FALSE}
here::i_am(
  "507.Rmd"
)

file_path<-here::here("data/pregnancy.csv")
data<-read.csv(file_path,header = TRUE)

data$mstatus[is.na(data$mstatus)]<-6
data$mstatus<-factor(data$mstatus,levels = c("0","1","2","3","4","5","6"))

data<-mice(data,m=5,method = "pmm",seed = 123)

data <- complete(data, 1)
data$mstatus<-factor(data$mstatus)
data$tpreg<-factor(data$tpreg)

data <- data %>%
  mutate(
    hb_dl = hb / 10,
    anemia = ifelse(hb_dl < 12, 1, 0),
    BMI = wt / (ht / 100)^2
  ) %>%
  dplyr::select(-ht, -wt, -hb, -hb_dl) 

data$crp <- log(data$crp+1)
data$sf <- log(data$sf+1)
data$sfo <- log(data$sfo+1)
data$si<-log(data$si+1)
data$vitb12<-log(data$vitb12+1)
data$age_yr<-log(data$age_yr+1)
data$gagewks<-log(data$gagewks+1)
data$BMI<-log(data$BMI+1)

data_anemia<-data%>%
  filter(anemia==1)
data_no_anemia<-data%>%
  filter(anemia==0)

times<-nrow(data_no_anemia)/nrow(data_anemia)

data_anemia_upsampled<-data_anemia[rep(1:nrow(data_anemia),each=ceiling(times)),]

data_balanced<-rbind(data_anemia_upsampled,data_no_anemia)
```

## Stepwise model

```{r,message=FALSE}

model<-glm(anemia ~ si+sf+sfo+vitb12+trimester+gagewks+age_yr+crp+mstatus+tpreg+BMI,family = binomial(link="logit"),data=data_balanced)

step.model =stepAIC(model, direction="both")
```


```{r}
model_step<-glm(anemia ~ si + sfo + vitb12 + trimester + gagewks + mstatus + BMI,family = binomial(link="logit"),data=data_balanced)
summary(model_step)

model_best<-glm(anemia ~ si + sfo + vitb12 + trimester + gagewks ,binomial(link="logit"),data=data_balanced)
summary(model_best)

model_bad<-glm(anemia~crp + mstatus + tpreg + BMI+age_yr+sf+trimester,family = binomial(link = "logit"),data=data_balanced)
summary(model_bad)

```

## Model evaluation

```{r}

# Calculate ROC objects
roc_step <- roc(response = data_balanced$anemia, predictor = fitted(model_step))
roc_best <- roc(response = data_balanced$anemia, predictor = fitted(model_best))
roc_bad <- roc(response = data_balanced$anemia, predictor = fitted(model_bad))

# Plotting all ROC curves together
plot(roc_step, main = "Comparison of ROC Curves", col = "red", print.auc = FALSE, print.auc.y = 0.2)
plot(roc_best, add = TRUE, col = "blue", print.auc = FALSE, print.auc.y = 0.4)
plot(roc_bad, add = TRUE, col = "green", print.auc = FALSE, print.auc.y = 0.6)

# Adding a legend to distinguish the models
legend("bottomright", legend = c("Model Step", "Model Best", "Model Bad"), 
       col = c("red", "blue", "green"), lwd = 2)

# Printing AUC scores separately
auc_step <- auc(roc_step)
auc_best <- auc(roc_best)
auc_bad <- auc(roc_bad)

print(paste("AUC for Model Step:", auc_step))
print(paste("AUC for Model Best:", auc_best))
print(paste("AUC for Model Bad:", auc_bad))

```

it seems like the stepwise model is still the best model.

```{r}
glmtoolbox::hltest(model_step)
glmtoolbox::hltest(model_best)
glmtoolbox::hltest(model_bad)
```

These models fit well.

## Alternative way

### Feature selection

```{r,echo=FALSE,include=FALSE}
data<-read.csv("C:/Users/pengfei lou/Desktop/homework/2024-1/applied linear regression/pregnancy.csv")

data$mstatus[is.na(data$mstatus)]<-6
#data$mstatus<-factor(data$mstatus,levels = c("0","1","2","3","4","5","6"))


data<-mice(data,m=5,method = "pmm",seed = 123)

data <- complete(data, 1)

data <- data %>%
  mutate(
    hb_dl = hb / 10,
    anemia = ifelse(hb_dl < 12, 1, 0),
    BMI = wt / (ht / 100)^2
  ) %>%
  dplyr::select(-ht, -wt, -hb, -hb_dl,-subjid) 

data$crp <- log(data$crp+1)
data$sf <- log(data$sf+1)
data$sfo <- log(data$sfo+1)
data$si<-log(data$si+1)
data$vitb12<-log(data$vitb12+1)
data$age_yr<-log(data$age_yr+1)
data$gagewks<-log(data$gagewks+1)
data$BMI<-log(data$BMI+1)

data_anemia<-data%>%
  filter(anemia==1)
data_no_anemia<-data%>%
  filter(anemia==0)

times<-nrow(data_no_anemia)/nrow(data_anemia)

data_anemia_upsampled<-data_anemia[rep(1:nrow(data_anemia),each=ceiling(times)),]

data_balanced<-rbind(data_anemia_upsampled,data_no_anemia)

```


```{r}

data_balanced$anemia <- factor(data_balanced$anemia, levels = c(0, 1))

set.seed(123) 
index <- createDataPartition(data_balanced$anemia, p = 0.8, list = FALSE)
train_data <- data_balanced[index, ]
test_data <- data_balanced[-index, ]

set.seed(7) # For reproducibility
outcomeVar <- "anemia"
predictorVars <- setdiff(names(train_data), outcomeVar)

# Define the control using a repeated cross-validation approach
control <- rfeControl(functions = rfFuncs, method = "cv", number = 10)

# Run the RFE algorithm
results <- rfe(train_data[, predictorVars], train_data[, outcomeVar],
               sizes = c(1:length(predictorVars)), rfeControl = control)

# Summarize the results
print(results)

# List the chosen features
chosen_predictors <- predictors(results)
print(chosen_predictors)

# Plot the RFE performance
plot(results, type = c("g", "o"))


#control <- trainControl(method="repeatedcv", number=10, repeats=3, search="grid")

# Train the model with variable selection using LASSO
#set.seed(7)
#model <- train(anemia ~ si+sf+sfo+vitb12+trimester+gagewks+age_yr+crp+mstatus+tpreg+BMI, data=train_data,method="glmnet",trControl=control,tuneLength=5,family="binomial")
# Check the results
#print(model)
#plot(varImp(model), top=11)
```

The max number of the model should be 9 and top 5 variables are  si, sfo, sf, vitb12, gagewks.

```{r}
set.seed(123)
index <- createDataPartition(data_balanced$anemia, p = 0.8, list = FALSE)
train_data <- data_balanced[index, ]
test_data <- data_balanced[-index, ]

# Ensure the outcome variable is a factor with appropriate levels
train_data$anemia <- factor(train_data$anemia, levels = c(0, 1))

# Fit a logistic regression model to the training data
model <- glm(anemia ~ si+sfo+sf+vitb12+gagewks+crp+age_yr+BMI+mstatus, 
             family = binomial(link = "logit"), data = train_data)

# Obtain predicted probabilities from the model
train_predictions_model <- predict(model, type = "response", newdata = train_data)

# Convert probabilities to binary class predictions
train_predicted_class <- ifelse(train_predictions_model > 0.5, 1, 0)
train_predicted_class <- factor(train_predicted_class, levels = c(0, 1))

# Generate a confusion matrix to see how well the model performs on the training data
confusion_matrix <- confusionMatrix(train_predicted_class, train_data$anemia)
print(confusion_matrix)


```


```{r}
set.seed(123)
# Ensure consistent use of dataset
index <- createDataPartition(data_balanced$anemia, p = 0.8, list = FALSE)
train_data <- data_balanced[index, ]
test_data <- data_balanced[-index, ]

# Ensure the outcome variable in the test data is a factor
test_data$anemia <- factor(test_data$anemia, levels = c(0, 1))

test_predictions_model <- predict(model, type = "response", newdata = test_data)

# Convert probabilities to binary class predictions and factorize
test_predicted_class <- ifelse(test_predictions_model > 0.5, 1, 0)
test_predicted_class <- factor(test_predicted_class, levels = c(0, 1))

# Evaluate the model performance on the test data
conf_matrix <- confusionMatrix(test_predicted_class, test_data$anemia)
print(conf_matrix)

```


```{r}
set.seed(123)
index <- createDataPartition(data_balanced$anemia, p = 0.8, list = FALSE)
train_data <- data_balanced[index, ]
test_data <- data_balanced[-index, ]

# Ensure the outcome variable is a factor with appropriate levels
train_data$anemia <- factor(train_data$anemia, levels = c(0, 1))

# Fit a logistic regression model to the training data
model_step<-glm(anemia ~ si + sfo + vitb12 + trimester + gagewks + mstatus + BMI,family = binomial(link="logit"),data=train_data)

# Obtain predicted probabilities from the model
train_predictions_step <- predict(model_step, type = "response", newdata = train_data)

# Convert probabilities to binary class predictions
train_predicted_class <- ifelse(train_predictions_step > 0.5, 1, 0)
train_predicted_class <- factor(train_predicted_class, levels = c(0, 1))

# Generate a confusion matrix to see how well the model performs on the training data
confusion_matrix <- confusionMatrix(train_predicted_class, train_data$anemia)
print(confusion_matrix)


```



```{r}
set.seed(123)
# Ensure consistent use of dataset
index <- createDataPartition(data_balanced$anemia, p = 0.8, list = FALSE)
train_data <- data_balanced[index, ]
test_data <- data_balanced[-index, ]

# Ensure the outcome variable in the test data is a factor
test_data$anemia <- factor(test_data$anemia, levels = c(0, 1))

# Assuming 'model' is already trained using train_data as shown in previous examples
test_predictions_step <- predict(model_step, type = "response", newdata = test_data)

# Convert probabilities to binary class predictions and factorize
test_predicted_class <- ifelse(test_predictions_step > 0.5, 1, 0)
test_predicted_class <- factor(test_predicted_class, levels = c(0, 1))

# Evaluate the model performance on the test data
conf_matrix <- confusionMatrix(test_predicted_class, test_data$anemia)
print(conf_matrix)
```

### AUC and ROC for train

```{r}
# ROC and AUC for model
roc_obj1 <- roc(response = train_data$anemia, predictor = train_predictions_model)
auc_value1 <- auc(roc_obj1)

# ROC and AUC for model_step
roc_obj2 <- roc(response = train_data$anemia, predictor = train_predictions_step)
auc_value2 <- auc(roc_obj2)

# Print AUC values
print(paste("AUC for Model :", auc_value1))
print(paste("AUC for Model_step:", auc_value2))

# Plot both ROC curves on the same plot
plot(roc_obj1, main="ROC Curve Comparison-train", col="red", print.auc=TRUE, print.auc.y=0.4)
plot(roc_obj2, add=TRUE, col="blue", print.auc=TRUE, print.auc.y=0.3)

# Adding a legend to distinguish the models
legend("bottomright", legend=c("Model ", "Model_step"), col=c("red", "blue"), lwd=2)
```

### AUC and ROC for test
```{r}

# ROC and AUC for model
roc_obj1 <- roc(response = test_data$anemia, predictor = test_predictions_model)
auc_value1 <- auc(roc_obj1)

# ROC and AUC for model_step
roc_obj2 <- roc(response = test_data$anemia, predictor = test_predictions_step)
auc_value2 <- auc(roc_obj2)

# Print AUC values
print(paste("AUC for Model :", auc_value1))
print(paste("AUC for Model_step:", auc_value2))

# Plot both ROC curves on the same plot
plot(roc_obj1, main="ROC Curve Comparison-test", col="red", print.auc=TRUE, print.auc.y=0.4)
plot(roc_obj2, add=TRUE, col="blue", print.auc=TRUE, print.auc.y=0.3)

# Adding a legend to distinguish the models
legend("bottomright", legend=c("Model ", "Model_step"), col=c("red", "blue"), lwd=2)

```

```{r}
glmtoolbox::hltest(model)

glmtoolbox::hltest(model_step)

```
The model-glm(anemia ~ si+sfo+sf+vitb12+gagewks+crp+age_yr+BMI+mstatus) turns out to be the best model for it has higer AUC socre in test performance and max number of variables in the model is 9. 
Both models fit well.










